Transitionin towards Continuous Delivery in the B2B Domain: A Case Study

### 2014-12-02 paper.tex
Renamed paper

% This is lnbip.tex the demonstration file of the LaTeX macro package for
% Lecture Notes in Business Information Processing from Springer-Verlag.
% It serves as a template for authors as well.
% version 1.0 for LaTeX2e
%
\documentclass[lnbip]{svmultln}
%
\usepackage{makeidx}  % allows for indexgeneration
% \makeindex          % be prepared for an author index

@@

\newsavebox\mybox
\newenvironment{aquote}[1]
  {\savebox\mybox{#1}\begin{quotation}}
  {\signed{\usebox\mybox}\end{quotation}}
%
\begin{document}
%
\mainmatter              % start of the contribution
%
\title{Title here}
%
\titlerunning{Abbreviated title asd}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Olli Rissanen\inst{1} %\and Roger Temam\inst{2}
%Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
}
%
\authorrunning{Olli Rissanen}   % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Olli Rissanen}
%
\institute{University of Helsinki, Helsinki, Finland,\\
\email{olli.rissanen@aalto.fi},\\ WWW home page:
\texttt{http://users/\homedir iekeland/web/welcome.html}
}

@@

\begin{abstract}        % give a summary of your study
Delivering more value to the customer is the goal of every software company. In modern software business, delivering value in real-time requires a company to utilize real-time deployment of software to expose features to users faster, and to shorten the feedback loop. This allows for faster reaction times, ensuring the development is focused on features providing real value. This study investigates practice known as continuous delivery andas means of providing value for the customers in real-time. Continuous delivery is a development practice where the software functionality is deployed continuously to customer environment. This process includes automated builds, automated testing and automated deployment. As a part of this paper, a case study is conducted in a medium-sized software company. The research objective is to analyze the challenges, benefits and organizational aspects of continuous delivery in the B2B domain. The data is collected from interviews conducted on members of two teams developing two different software products. The results suggest that technical challenges are only one part of the challenges a company encounters in this transition. For continuous delivery, the company must also address challenges related to the customer and procedures. The core challenges are caused by having multiple customers with diverse environments and unique properties, whose business depends on the software product. Some customers also require to perform manual acceptance testing, which slows down production deployments. The benefits found from continuous delivery support the case company in solving many of its business problems. The company can expose the software functionality to the customers from an earlier stage, and guide the product development by utilizing feedback and data instead of opinions.
%                         please supply keywords within your abstract
\keywords {Continuous delivery, Development process, B2B}
\end{abstract}
%
\section{Introduction}

@@

\noindent Software development practices and product characteristics vary based on the domain and delivery model. Typical B2C applications are hosted as Software as a Service (SaaS) applications, and accessed by users via a web browser. In the B2B domain, applications installed to customer environments are very common. Running experiments on customer environments also requires a new deployment of the software product, possibly causing downtime and changing some functionality of the software. The purpose of this research question is to identify the challenges faced in applying these practices in the B2B environment. The research question is answered by conducting interviews to discover the current development process and its challenges in the case company, and using these findings and the available literature on continuous delivery and continuous experimentation to map the initial set of challenges these approaches will face in the case company. The available literature is used to provide a thorough understanding of continuous delivery and continuous experimentation as a whole, so that challenges can be identified in all aspects of the practices.
%Question: contractual issues? 
%Question: selecting the correct OEC in b2b is harder. for example revenue. 
%question: in B2C people develop their own product, but in B2B the feature ideas might come only from customer.
\newline

@@

\noindent To rationalize the decision to adopt continuous delivery and experimentation in a company, the actual benefits to the business have to be identified. This research question aims to identify clear objectives for what is achieved by adopting continuous delivery and continuous experimentation. Sections of the interview aim to identify the current perceived problems of the case company related to deployment, product development, collecting feedback and guiding the development process. These problems are then compared to the benefits of these approaches found from the literature.
\newline

@@

%release cycles
In an agile process software release is done in periodic intervals \cite{cockburn2002agile}. Compared to waterfall model it introduces multiple releases throughout the development. Continuous delivery, on the other hand, attemps to keep the software ready for release at all times during development process \cite{cdbook}. Instead of stopping the development process and creating a build as in an agile process, the software is continuously deployed to customer environment. This doesn't mean that the development cycles in continuous delivery are shorter, but that the development is done in a way that makes the software always ready for release.
\begin{figure}[h]
  \centering
  \includegraphics[width=5.0in]{rtvd.jpg}
  \caption{Continuous integration, delivery and deployment.}
  \label{fig2}
\end{figure}
%delivery vs deployment
Continuous delivery differs from continuous deployment. Refer to Fig. \ref{fig2} for a visual representation of differences in continuous integration, delivery and deployment. Both continuous delivery and continuous deployment include automated deployment to a staging environment. Continuous deployment includes deployment to a production environment, while in continuous delivery the deployment to a production environment is done manually. The purpose of continuous delivery is to prove that every build is deployable \cite{cdbook}. While it necessarily doesn't mean that teams release often, keeping the software in a state where a release can be made instantly is often seen beneficial.

%Barriers in CI->CD transition: stairway
Holmström Olsson et al. research the transition phase from continuous integration to continuous delivery, and identify barriers that companies need to overcome to achieve the transition \cite{olsson2012climbing}. One such barrier is the custom configuration at customer sites. Maintaining customized solutions and local configurations alonside the standard configurations creates issues. The second barrier is the internal verification loop, that has to be shortened not only to develop features faster but also to deploy fast. Finally, the lack of transparency and getting an overview of the status of development projects is seen as a barrier.

@@

@@

The purpose of this study is to seek ways to improve the product development process of two software products of the case company in the B2B domain, utilizing a practice known as continuous delivery. This includes analysing the challenges and benefits of this practice. Existing documented implementations of experimentational systems are primarily executed in the B2C domain, often with a Software as a Service (SaaS) product. An example of this is the Microsoft EXP platform \cite{ep}. The focus of this study is in the B2B domain, with two software products that are not used as SaaS and are developed directly to other companies acting as customers.

\bigskip
\noindent \textbf{Research objective}
To analyze the software development process transition towards shorter feedback cycle and data-driven decisions from the perspective of the case company, with respect to continuous delivery and continuous experimentation practices.

\subsection{Research method}
%Methods
The primary source of information in this research are semi-structured interviews \cite{runeson2009guidelines} performed within the CDM and Dialog teams. The interview consists of pre-defined themes as follows: (1) current development process, (2) current deployment process, (3) current interaction with customers, (4) software product and (5) future ways with continuous delivery and continuous experimentation. Data is also collected through the product description documents and development process documents to verifying and supplementing the interview data. Data triangulation is implemented by interviewing multiple individuals in different positions of the teams. Methodological triangulation is implemented by collecting documentary data regarding the development process from internal documents, and by including observations made by the author in the analysis.

@@

%Data analysis
In this case study the data analysis is based on template analysis, which is a way of thematically analysing qualitative data \cite{king1998template}. The initial template was first formed by exploring the qualitative data for three themes: development process, deployment of software and experimentation. Through multiple iterations of the data, multiple subthemes were then added to the three existing themes by further coding the data. Attention was paid to different roles of the interviewees.
\subsection{Execution}

@@

This section addresses the RQ1 (i.e. What are the B2B specific challenges of continuous delivery and continuous experimentation?). The challenges regarding continuous delivery are analyzed in three areas: technical, procedural and customer. Technical aspect includes the environmental challenges, configural challenges and other challenges related to the software product and its usage. Procedural aspect includes the challenges regarding the development process of software. Customer aspect consists of the customer interaction process and customer commitment.

@@

The study investigates extending the development process towards continuous delivery and continuous experimentation to advance the company towards real-time value delivery. The research questions address the challenges faced in this transition, the benefits found from these practices and systematical application of continuous experimentation in the B2B domain.

@@

%to theory
The findings are in align with and could be considered as extending some of the theoretical contributions by Olsson et al. \cite{olsson2012climbing}, who researched the transition towards continuous delivery and an experimentational system. Identifying that transitioning towards continuous delivery requires a company to address issues in multiple aspects of the company also benefits companies in practice.

@@

\paragraph{Notes and Comments.}
asd

>>>>

% This is lnbip.tex the demonstration file of the LaTeX macro package for
% Lecture Notes in Business Information Processing from Springer-Verlag.
% It serves as a template for authors as well.
% version 1.0 for LaTeX2e
%
\documentclass[lnbip]{svmultln}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{graphicx}
%\usepackage{graphics} 
% \makeindex          % be prepared for an author index

@@

\newsavebox\mybox
\newenvironment{aquote}[1]
  {\savebox\mybox{#1}\begin{quotation}}
  {\signed{\usebox\mybox}\end{quotation}}
%
\begin{document}
%
\mainmatter              % start of the contribution
%
\title{Transitionin towards Continuous Delivery in the B2B Domain: A Case Study}
%
\titlerunning{Abbreviated title}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Olli Rissanen \and J{\"u}rgen M{\"u}nch
%Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
}
%
\authorrunning{Olli Rissanen}   % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Olli Rissanen}
%
\institute{University of Helsinki, Helsinki, Finland,\\
\email{olli.rissanen@aalto.fi}
}

@@

\begin{abstract}        % give a summary of your study
Delivering more value to the customer is the goal of every software company. In modern software business, delivering value in real-time requires a company to utilize real-time deployment of software to expose features to users faster, and to shorten the feedback loop. This allows for faster reaction times, ensuring the development is focused on features providing real value. This study investigates a practice known as continuous delivery as means of providing value for the customers in real-time. Continuous delivery is a development practice where the software functionality is deployed continuously to customer environment. This process includes automated builds, automated testing and automated deployment. As a part of this paper, a case study is conducted in a medium-sized software company. The research objective is to analyze the challenges, benefits and organizational aspects of continuous delivery in the B2B domain. The data is collected from interviews conducted on members of two teams developing two different software products. The results suggest that technical challenges are only one part of the challenges a company encounters in this transition. For continuous delivery, the company must also address challenges related to the customer and procedures. The core challenges are caused by having multiple customers with diverse environments and unique properties, whose business depends on the software product. Some customers also require to perform manual acceptance testing, which slows down production deployments. The benefits found from continuous delivery support the case company in solving many of its business problems. The company can expose the software functionality to the customers from an earlier stage, and better utilize customer feedback.
%                         please supply keywords within your abstract
\keywords {Continuous delivery, Development process, B2B}
\end{abstract}
%
\section{Introduction}

@@

\noindent Software development practices and product characteristics vary based on the domain and delivery model. Typical B2C applications are hosted as Software as a Service (SaaS) applications, and accessed by users via a web browser. In the B2B domain, applications installed to customer environments are very common. The purpose of this research question is to identify the challenges faced in applying continuous delivery in the B2B environment. The research question is answered by conducting interviews to discover the current development process and its challenges in the case company, and using these findings and the available literature on continuous delivery to map the initial set of challenges these approaches will encounter in the case company. The available literature is used to provide a thorough understanding of continuous delivery as a whole, so that challenges can be identified in all aspects of the practice.
%Question: contractual issues? 
%Question: selecting the correct OEC in b2b is harder. for example revenue. 
%question: in B2C people develop their own product, but in B2B the feature ideas might come only from customer.
\newline

@@

\noindent To rationalize the decision to adopt continuous delivery in a company, the actual benefits to the business have to be identified. This research question aims to identify clear objectives for what is achieved by adopting continuous delivery. Sections of the interview aim to identify the current perceived problems of the case company related to deployment, product development, collecting feedback and guiding the development process. These problems are then compared to the benefits of the approach found from the literature.
\newline

@@

%release cycles
In an agile process software release is done in periodic intervals \cite{cockburn2002agile}. Compared to waterfall model it introduces multiple releases throughout the development. Continuous delivery, on the other hand, attemps to keep the software ready for release at all times during development process \cite{cdbook}. Instead of stopping the development process and creating a build as in an agile process, the software is continuously deployed to customer environment. This doesn't mean that the development cycles in continuous delivery are shorter, but that the development is done in a way that makes the software always ready for release.
\begin{figure}[h]
  \centering
  \includegraphics[width=5.0in]{rtvd.jpg}
  \caption{Continuous integration, delivery and deployment.}
  \label{fig2}
\end{figure}

%\begin{figure}
% use the \includegraphics command from the graphics package
% (N.B. put \usepackage{graphics} into your preamble then)
% for inclusion of your images. A sample call could read:
% \includegraphics{myimage}
%\vspace{2.5cm}
%\caption{This is the caption of the figure displaying a white eagle and
%a white horse on a snow field}
%\end{figure}
%delivery vs deployment
Continuous delivery differs from continuous deployment. Refer to Fig. \ref{fig2} for a visual representation of differences in continuous integration, delivery and deployment. Both continuous delivery and continuous deployment include automated deployment to a staging environment. Continuous deployment includes deployment to a production environment, while in continuous delivery the deployment to a production environment is done manually. The purpose of continuous delivery is to prove that every build is deployable \cite{cdbook}. While it necessarily doesn't mean that teams release often, keeping the software in a state where a release can be made instantly is often seen beneficial.

@@

%Barriers in CI->CD transition: stairway
Holmström Olsson et al. research the transition phase from continuous integration to continuous delivery, and identify barriers that companies need to overcome to achieve the transition \cite{olsson2012climbing}. One such barrier is the custom configuration at customer sites. Maintaining customized solutions and local configurations alongside the standard configurations creates issues. The second barrier is the internal verification loop, that has to be shortened not only to develop features faster but also to deploy fast. Finally, the lack of transparency and getting an overview of the status of development projects is seen as a barrier.

@@

The purpose of this study is to seek ways to improve the product development process of two software products of the case company in the B2B domain, utilizing a practice known as continuous delivery. This includes analysing the challenges and benefits of this practice. Existing documented implementations of continuous delivery are primarily executed in the B2C domain, often with a Software as a Service (SaaS) product. The focus of this study is in the B2B domain, with two software products that are not used as SaaS and are developed directly to other companies acting as customers. The characteristics of this domain are typically longer release cycles and customer owned environments.

\bigskip
\noindent \textbf{Research objective}
To analyze the software development process transition towards shorter feedback cycle and data-driven decisions from the perspective of the case company in the B2B domain, with respect to continuous delivery.

\subsection{Research method}
%Methods
The primary source of information in this research are semi-structured interviews \cite{runeson2009guidelines} performed within the CDM and Dialog teams. The interview consists of pre-defined themes focusing on current development process, current deployment process, current interaction with customers, the software products and future ways with continuous delivery. Data is also collected through the product description documents and development process documents to verifying and supplementing the interview data. Data triangulation is implemented by interviewing multiple individuals in different positions of the teams. Methodological triangulation is implemented by collecting documentary data regarding the development process from internal documents, and by including observations made by the author in the analysis.

@@

%Data analysis
In this case study the data analysis is based on template analysis, which is a way of thematically analysing qualitative data \cite{king1998template}. The initial template was first formed by exploring the qualitative data for two themes: development process and deployment of software. Through multiple iterations of the data, multiple subthemes were then added to the two existing themes by further coding the data. Attention was paid to different roles of the interviewees.
\subsection{Execution}

@@

This section addresses the RQ1 (i.e. What are the B2B specific challenges of continuous delivery?). The challenges regarding continuous delivery are analyzed in three areas: technical, procedural and customer. Technical aspect includes the environmental challenges, configural challenges and other challenges related to the software product and its usage. Procedural aspect includes the challenges regarding the development process of software. Customer aspect consists of the customer interaction process and customer commitment.

@@

The study investigates extending the development process towards continuous delivery to advance the company towards real-time value delivery. The research questions address the challenges faced in this transition, and the benefits found from this practice.

@@

%to theory
The findings are in align with and could be considered as extending some of the theoretical contributions by Olsson et al. \cite{olsson2012climbing}, who researched the transition towards continuous delivery. Identifying that transitioning towards continuous delivery requires a company to address issues in multiple aspects of the company also benefits companies in practice.

@@
